---
title: "Find and plot correlated vote results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{test}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### What Other Votes Correlate with the 2014 vote on immigration (mei)?

For the sake of illustration we are interested in the result of the 2014 vote on immigration for Swiss municipalities. The initiative was voted upon on 9th of February 2014. There are different ways to find this exact vote. One of them is using the information in the title. An other solution would be to look up the `id` provided by the FSO. 

First, we invoke the necessary packages and use the function `get_nationalvotes` to access the data. We further specify the unit of analysis as well as the range 

unique(federalvotes$name[grep("Massen", federalvotes$name)])

```{r echo=TRUE, warning=FALSE, message=FALSE}
# installation from CRAN (stable)
# install.packages("swissdd")
# install.packages("dplyr")

# installation from github (ongoing updates)
# devtools::install_github("politanch/swissdd")

library(swissdd)
library(dplyr)
library(ggplot2)
library(tidyr)

#get data from API for the 2010-2019 (feb) 
federalvotes <- get_nationalvotes(geolevel = "municipality", 
                                  from_date = "2010-03-07", 
                                  to_date = "2019-02-10")

#get correlations for votes on municipal level with mei
simvotes <- similar_votes(federalvotes, id=5800, from=.4, to=.6)
simvotes

#extract names of correlated votes
ballotnames <- federalvotes %>%
  dplyr::select(name, id, mun_id)%>%
  filter(id%in%c(5800, simvotes[1,1]))%>%
  distinct()

#subset for correlated votes
corrvotes <- federalvotes %>% 
  filter(id%in%c(5800, simvotes[1,1]))%>%
  dplyr::select(id, jaStimmenInProzent, mun_id)%>%
  mutate(id=as.character(id))

#plot
spread(corrvotes , id, jaStimmenInProzent)%>%
  ggplot(aes(y=`5800`, x=`5990`))+
  geom_point()+
  scale_y_continuous(limits=c(0,100))+
  scale_x_continuous(limits=c(0,100))+
  geom_abline(intercept = 0, slope=1,  size=.1)+
  geom_smooth(method="lm", size=.1, color="orange")

```

## Buy us a coffee

`swissdd` is an open source project, developed by us. If `swissdd` is useful to you, consider buying us a coffee, a beer or making a monthly donation so we can keep building great free software.

<link href="https://fonts.googleapis.com/css?family=Arial" rel="stylesheet"><a class="bmc-button" target="_blank" href="https://www.buymeacoffee.com/politanch"><img src="https://cdn.buymeacoffee.com/buttons/bmc-new-btn-logo.svg" alt="Buy us a coffee"><span style="margin-left:15px;font-size:19px !important;">Buy us a coffee</span></a>