---
title: "Find and plot correlated vote results"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### What other votes correlate with the 2014 vote on immigration (mei)?

For the sake of illustration we are interested in the result of the 2014 vote on immigration for Swiss municipalities. The initiative was voted upon on 9th of February 2014. There are different ways to find this exact vote. One of them is using the information in the title. An other solution would be to look up the `id` provided by the FSO. 

<<<<<<< HEAD
=======
`unique(federalvotes$name[grep("Massen", federalvotes$name)])`

>>>>>>> master
First, we invoke the necessary packages and use the function `get_nationalvotes` to access the data. We further specify the unit of analysis as well as the range.


```{r echo=TRUE, warning=FALSE, message=FALSE}
# installation from CRAN (stable)
# install.packages("swissdd")
# install.packages("dplyr")

# installation from github (ongoing updates)
# devtools::install_github("politanch/swissdd")

library(swissdd)
library(dplyr)
library(ggplot2)
library(tidyr)

#get results of all votes between 2010-2019
federalvotes <- get_nationalvotes(geolevel = "municipality", 
                                  from_date = "2010-03-07", 
                                  to_date = "2020-09-27")

#get correlations for votes on municipal level with mei
simvotes <- similar_votes(federalvotes, id=5800, from=.4, to=.6)
simvotes

#extract names of correlated votes
ballotnames <- federalvotes %>%
  dplyr::select(name, id, mun_id)%>%
  filter(id%in%c(5800, simvotes[2,1]))%>%
<<<<<<< HEAD
  distinct(name)
=======
  distinct()
>>>>>>> master

#subset for correlated votes
corrvotes <- federalvotes %>% 
  filter(id%in%c(5800, simvotes[2,1]))%>%
  dplyr::select(id, jaStimmenInProzent, mun_id)%>%
  mutate(id=as.character(id))

#plot
<<<<<<< HEAD
corrvotes%>%
  pivot_wider(names_from="id", values_from="jaStimmenInProzent")%>%
=======
spread(corrvotes , id, jaStimmenInProzent)%>%
>>>>>>> master
  ggplot(aes(y=`5800`, x=`5710`))+
  geom_point()+
  scale_y_continuous(limits=c(0,100))+
  scale_x_continuous(limits=c(0,100))+
  geom_abline(intercept = 0, slope=1,  size=.1)+
  geom_smooth(method="lm", size=.1, color="orange")+
  labs(y="MEI, Yes Shares", x="Amendments to the Asylum Law, Yes Shares", caption="politan.ch")+
  theme_bw()

```




### What Other Votes Correlate with the 2020 vote on limiting immigration (BGI)?


```{r echo=TRUE, warning=FALSE, message=FALSE}

#get correlations for votes on municipal level with mei
simvotes <- similar_votes(federalvotes, id=6310, from=.3, to=1)
simvotes

#extract names of correlated votes
ballotnames <- federalvotes %>%
  dplyr::select(name, id, mun_id)%>%
  filter(id%in%c(6310, simvotes[1,1]))%>%
  distinct(name)

#subset for correlated votes
corrvotes <- federalvotes %>% 
  filter(id%in%c(6310, simvotes[1,1]))%>%
  dplyr::select(id, jaStimmenInProzent, mun_id)%>%
  mutate(id=as.character(id))

#plot
corrvotes%>%
  pivot_wider(names_from="id", values_from="jaStimmenInProzent")%>%
  ggplot(aes(y=`6310`, x=`5800`))+
  geom_point(alpha=.2)+
  scale_y_continuous(limits=c(0,100))+
  scale_x_continuous(limits=c(0,100))+
  geom_abline(intercept = 0, slope=1,  size=.1)+
  geom_smooth(method="lm", size=.3, color="orange")+
  labs(y="BGI, Yes Shares", x="MEI, Yes Shares", caption="politan.ch")+
  theme_bw()









