---
title: "Plot vote shares"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This functionality has been kindly added by [David Zumbach](https://twitter.com/davidzumbach){target="_blank"}.

The function `get_geodata` allows users to download geographic information for different administrative units in Switzerland. The function returns a *simple feature* `data.frame`. Users should specify the level of administrative units using the argument `geolevel`. The following options are available:

* `national` to download administrative boundaries of Switzerland
* `cantonal` to download administrative boundaries of cantons
* `municipality` to download administrative boundaries of municipalities/communities
* `zh_counting_districts` to download administrative boundaries of *Zählkreise* for the City of Zurich
* `lakes` to download boundaries of major lakes

An illustration of how one could utilize the functionionality of `swissdd` to visualize vote outcomes for any national vote is given here.

### Plot voteshares "by hand"

Producing a map "by hand" allows for more flexibility. For example, you could simply download the geo-information and plot something entirely else. If you never ever want to produce a plot by hand, skip to the next sectionn.


```{r echo=TRUE, warning=FALSE, message=FALSE}
# installation from CRAN (stable)
# install.packages("swissdd")
# install.packages("dplyr")
# installation from github (ongoing updates)
# devtools::install_github("politanch/swissdd", ref="dev")

library(swissdd)
packageVersion("swissdd")
library(dplyr)
library(ggplot2)
library(sf)
library(viridis)

#download geo information
geo_canton <- swissdd::get_geodata(geolevel = "canton")
geo_canton$canton_id <- as.numeric(geo_canton$canton_id)

# download data from API on the vote calles «Swiss coalition for Corporate justice SCCJ»
kovi_nat <- swissdd::get_nationalvotes(votedates="2020-11-29", geolevel = "canton")%>%
  dplyr::filter(id == 6360)%>%
  dplyr::select(canton_id, jaStimmenInProzent)%>%
  mutate(canton_id=as.numeric(canton_id))
```

Combine the two `data.frames`.
```{r echo=TRUE, warning=FALSE, message=FALSE}
can_df <- left_join(geo_canton, kovi_nat, by="canton_id")
```

Plot the whole thing accordingly.

```{r echo=TRUE, warning=FALSE, message=FALSE}
ggplot(can_df)+
  geom_sf(aes(fill=jaStimmenInProzent), color="white", size=.1)
```

With this you can start prettifiyng the map according to your liking. Wee add the lakes as well here:

```{r echo=TRUE, warning=FALSE, message=FALSE}
lakes <- swissdd::get_geodata(geolevel = "lakes")
ggplot(can_df)+
  geom_sf(aes(fill=jaStimmenInProzent), color="white", size=.1)+
  geom_sf(data=lakes, fill="aliceblue", color=NA)+
  scale_fill_viridis(option="D", 
                     name = "Yes-Share",
                     guide=guide_colorbar(title.position="top", 
                                          direction = "horizontal",
                                          barheight = unit(2, units = "mm"),
                                          barwidth = unit(50, units = "mm"),
                                          title.hjust =0.5))+
  theme_void()+
  theme(legend.position="bottom")+
  labs(title="Result for Vote on Corporate Justice SCCJ", subtitle="November 29, 2020",
       caption="data: FSO/{swissdd}")
```


### Plot voteshares with built-in function

If for some reason you don't want to create your map from scratch you could also rely on the function `plot_nationalvotes()` which allows you to quickly plot turnout rates or yes-shares. In order to run this function you have to specify the administrative level as well as the official identification number of the vote you're interested in. 

```{r echo=TRUE, warning=FALSE, message=FALSE}
plot_nationalvotes(vote_id=6360, geolevel="canton")
```



