---
title: "Example: Plot Vote Shares"
output: github_document 
date: "Last Updated: `r format(Sys.time(), '%d, %B, %Y at %H:%M')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### How to Use Swissvotes Data

First, you have to make sure that your machine is connected to the internet. 

```{r echo=TRUE, warning=FALSE, message=FALSE}
# installation from CRAN (stable)
# install.packages("swissdd")
# install.packages("dplyr")

# installation from github (ongoing updates)
# devtools::install_github("politanch/swissdd")

Sys.setlocale('LC_ALL','C')

library(swissdd)
library(dplyr)

```


By default, function extracts only the database provided by https://swissvotes.ch. However, you can specify that you want the codebook as well (or just the codebook for that matter). Additionally, there is a specification to save the citation. If you work with Swissvotes data, we recommend to save the citation.

If you don't specify anything the function will be executed with the following specifications.

```{r echo=TRUE, warning=FALSE, message=FALSE, eval=FALSE}
#default
swissvotesDB <- get_swissvotes(DB=T, savecitation=F, codebook=F)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#default
swissvotesDB <- get_swissvotes()
```

If you use the parameter `codebook=T` the function will download the data and direct your browser automatically to the codebook. If you're only interested in the codebook (you accessed the data already, for example), set the parameter `DB` to `FALSE`. 

Since we loaded the data already, there is no need to download it again.

```{r echo=TRUE, warning=FALSE, message=FALSE, eval=FALSE}

get_swissvotes(DB=F, codebook=T)

```

For illustration, we are interested in the upcoming vote on affordable housing (date of vote 9th of February, 2020). The variable `titel_kurz_d` or `titel_kurz_f` contains the title of a vote (shorter version than the official title `titel_off_d` or `titel_off_f`). Please note that the letter `d` indicates the German title, while the letter `f` gives you the French title. Since we don't know too much about the initiative we can read about it as well. First, we find the position in the data and then use the hyperlink provided to read about it.

```{r echo=TRUE, warning=FALSE, message=FALSE, eval=TRUE}

#get position in dataset
pos_ah <- grep("bezahlbare Wohnungen", swissvotesDB$titel_kurz_d)

#extract url that directs you to more information about the vote
mei_url <- swissvotesDB$anneepolitique[pos_ah]

#mei_url
#"https://anneepolitique.swiss/prozesse/56996-volksinitiative-mehr-bezahlbare-wohnungen"
```

```{r echo=TRUE, warning=FALSE, message=FALSE, eval=FALSE}
#access url
browseURL(mei_url)
```





